Synthesizing Symmetric Lenses

### 2018-07-10 papers/SymmetricLenses/main.tex
fixed minor typos
in progress edits to intro

\noindent
In a big-data world, one will often find similar data stored in a variety of
different formats. For instance, electronic calendars come in iCalendar,
vCalendar and h-event formats~\cite{calendar-formats}; US taxation data can be
transmitted via Tax XML (used by the US government for e-filing) or TXF (used by
TurboTax)~\cite{tax-formats}; and sewing machines can read embroidery files
transmitted as PES formats, DST formats or several
others~\cite{embroidery-formats}.

One way to interconvert between these varied data formats is to use a
\emph{lens}~\cite{?}---a bi-directional
program that can transform data represented in a format $S$ to a format $T$,
and vice versa.  There are a wide variety of different kinds of lenses, but well-defined lens
languages typically come with guarantees that help ensure data is
not lost or corrupted as it is transformed back and forth between different representations.

Lens languages, like any domain-specific language, also cut down the space of
possible programs one can write, and the smaller the space of programs,
the easier it is to find the program one is looking for.  
%This feature can be a negative, if the desired transformation
%falls outside of the scope provided by the language, but can also be positive, as complex
%properties (like invertability) can be enforced automatically.
%Another advantage of reducing the
%set of programs is that doing so can facilitate program synthesis, which is essentially
%a search through the space of possible computations that can be expressed as a lens.
Recent research~\cite{optician, maina+:quotient-synthesis} has exploited this
fact by demonstrating that it is possible to synthesize relatively large,
complex lenses from specifications. In particular, given types that describe
source and target formats, the bijective synthesis algorithm presented in this
research will find bijective lenses
% ~\cite{optician}
for files in ad-hoc formats, such as those housing linux application
configurations. Oftentimes, these algorithms find such lenses in less than a
second, and without the aid of any examples at all, with just types describing
the source and target formats.
% \dpw{I would cite the ICFP paper (cite string:
%   maina+:quotient-synthesis) as to appear but I'm not sure how to do that
%   without possibly violating the spirit of double-blind reviewing. ie, if it
%   hasn't appeared yet, how could we know the content if we aren't the authors?
%   Mentioning it in the 3rd person does not seem to suffice ...} \ksf{The ICFP
%   2018 website lists the paper; I think it is okay to cite it. The reviewer can
%   assume we contacted the authors for a copy. We could ask Stephanie for a
%   ruling if we want to be sure. One of the principles of double blind as adopted
%   by SIGPLAN is not to reduce the quality of the paper; omitting this citation
%   would have that affect.}\bcp{+1}\saz{Stephanie says we should cite it in the
%   third person, even if that's a bit awkward, but otherwise citing it is OK.}

One of reasons these algorithms are so effective, even relative to successful
synthesis projects in other domains, is because unless the same kind of data
appears in a format many times, there simply are not very many different
bijections between related data formats. Hence, if the synthesis algorithm can
find any bijection at all, then it is very likely to be the one the user wants.
However, there is also a price for narrowing the synthesis domain so
dramatically: Many real data formats are almost, but not quite, bijections. More
often two related data formats have some data in common, but also contain some
differences. For instance, two related tables might share four of five columns,
but not the last. Alternately, one ad hoc configuration file might host some
format-specific meta data, such as a date, a time or reference number, whereas
the other does not. Indeed, all of our benchmarks taken from Flash
Fill~\cite{flashfill} and all of our benchmarks that synchronize between two
ad-hoc file formats. Consequently, past tools that search for strict bijections
will rarely be applicable in practice.

>>>>

\noindent
In a big-data world, similar data is often stored in a variety of
formats. For instance, electronic calendars come in iCalendar,
vCalendar and h-event formats~\cite{calendar-formats}; US taxation data can be
transmitted via Tax XML (used by the US government for e-filing) or TXF (used by
TurboTax)~\cite{tax-formats}; and sewing machines can read embroidery files
transmitted as PES formats, DST formats or several
others~\cite{embroidery-formats}.

One way to interconvert between these varied data formats is to use a
\emph{lens}~\cite{Focal2005}---a bi-directional
program that can transform data represented in a format $S$ to a format $T$,
and vice versa.  There are many different kinds of lenses, but well-defined lens
languages typically provide guarantees that help ensure data is
not lost or corrupted as it is transformed back and forth between different representations.

While lenses facilitate writing principled data
transformations, \textit{synthesizing} lenses 
can make the task even easier.  Lens languages are a ripe target for
synthesis because lens languages, like many domain-specific languages,
reduce the space of possible programs, making search-based synthesis easier. 
%This feature can be a negative, if the desired transformation
%falls outside of the scope provided by the language, but can also be positive, as complex
%properties (like invertability) can be enforced automatically.
%Another advantage of reducing the
%set of programs is that doing so can facilitate program synthesis, which is essentially
%a search through the space of possible computations that can be expressed as a lens.
Recent research~\cite{optician, maina+:quotient-synthesis} has exploited this
observation by demonstrating that it is possible to synthesize relatively large,
complex lenses from specifications. In particular, given types that describe
source and target formats and optional examples of corresponding data, the
bijective synthesis algorithm presented in these 
papers will find bijective lenses to convert between data
% ~\cite{optician}
in ad hoc formats, such as linux 
configuration files that represent the same data in different ways. 
Often, these algorithms can find an appropriate lens
in less than a second without requiring any examples.
%, with just types describing the source and target formats.
% \dpw{I would cite the ICFP paper (cite string:
%   maina+:quotient-synthesis) as to appear but I'm not sure how to do that
%   without possibly violating the spirit of double-blind reviewing. ie, if it
%   hasn't appeared yet, how could we know the content if we aren't the authors?
%   Mentioning it in the 3rd person does not seem to suffice ...} \ksf{The ICFP
%   2018 website lists the paper; I think it is okay to cite it. The reviewer can
%   assume we contacted the authors for a copy. We could ask Stephanie for a
%   ruling if we want to be sure. One of the principles of double blind as adopted
%   by SIGPLAN is not to reduce the quality of the paper; omitting this citation
%   would have that affect.}\bcp{+1}\saz{Stephanie says we should cite it in the
%   third person, even if that's a bit awkward, but otherwise citing it is OK.}
One reason that bijective lens synthesis is so effective, even relative to successful
synthesis projects in other domains, is that there are simply not very
many different bijections between related data formats (unless the
same kind of data appears in the two formats many times). 
Hence, if the synthesis algorithm can
find any bijection at all, then it is very likely to be the one the
user wants.

However, there is a price for narrowing the synthesis domain so
dramatically: Many real data formats are almost, but not quite,
bijections. More often two related data formats have some data in
common, but also contain some differences. For instance, two related
tables might share four of five columns, but each have a different
fifth column. Alternately, one ad hoc configuration file might include
some file-specific meta data, such as a date, a time or reference
number, whereas the other does not. Indeed, all of our benchmarks
taken from Flash Fill~\cite{flashfill} and all of our benchmarks that
synchronize between two ad hoc file formats described in
Section \ref{sec:evaluation} \ksf{is this right?} have this 
characteristic. Consequently, past tools that search for strict
bijections will rarely be applicable in practice.
